% Appendix A
\chapter{Potential Conferences} % Main appendix title

\label{AppendixA} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}

\lhead{Appendix A. \emph{Potential Conferences}} % Change X to a consecutive letter; this is for the header on each page - perhaps a shortened title

The scientific paper may be published in available conferences where the submission deadlines are in accordance with the working plan for the second stage of this work. The following list shows some potential candidates:
\begin{itemize}
	\item \textbf{Association for Computational Linguistics \href{http://acl2016.org}{(ACL)}} - The ACL conference will be held from August 7 to 12, 2016 in Berlin, Germany. Important dates are February 29, 2016 for short papers submissions and March 18, 2016 for long papers submissions.

	\item \textbf{European Conference on Machine Learning and Principles and Practice of Knowledge Discovery \href{http://www.ecmlpkdd2016.org}{(ECML-PKDD)}} - The ECML-PKDD conference will be held from September 19 to 23, 2016 in Riva del Garda, Italy. Important dates are April 1, 2016 for abstract submissions and April 4, 2016 for papers submissions.

    \item \textbf{European Conference on Artificial Intelligence \href{http://www.ecai2016.org}{(ECAI)}} - The ECAI conference will be held from August 29 to September 2, 2016 in The Hague, The Netherlands. An important date is April 15, 2016 for papers submissions.
    
   % \item \textbf{Workshop on Language Technologies and Applications \href{https://fedcsis.org/2016/lta}{(LTA)}} - The LTA workshop will be held from September 11 to 14, 2016 in Gdansk, Poland. An important date is April 18, 2016 for papers submissions.
    
    \item \textbf{Conference on Computational Natural Language Learning \href{http://www.conll.org/cfp-2016}{(CoNLL)}} - The CoNLL conference will be held from August 11 to 12, 2016 in Berlin, Germany. An important date is May 6, 2016 for papers submissions.
    
    \item \textbf{Conference on Intelligent Text Processing and Computational Linguistics \href{http://www.cicling.org/2016}{(CICLing)}} - The CICLing conference will be held from April 3 to 9, 2016 in Konya, Turkey. Important dates are yet to be announced.
    
    \item \textbf{Conference on Computational Linguistics \href{http://coling2016.anlp.jp}{(COLING)}} - The COLING conference will be held from December 11 to 16, 2016 in Osaka, Japan. An important date is July 15, 2016 for papers submissions.
    
\end{itemize}
%

